---
title: "Lab 02 - Plastic waste"
author: Ben Waggener
date: 01/21/2025
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

Here is the plastic-waste-continent data plotted as a frequency graph 

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)
```

Here is the dataset filtered so that it only shows countries with plastic waste per capita that is greater than 3.5kg per person
```{r filtered_plasticwastecontinent}

plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```
This shows that the outlier from the previous graph is Trinidad and Tobago. This means that this country has a much higher rate of plastic waste per capita compared to the other countries in this dataset. Quickly looking online, it seems that this is not a "data error" and that much of this is caused by "inadequate waste management" (according to Forbes)


Below shows each continent as a density plot.

```{r desnsityplotbycontinent}

ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent
  )
) +
  geom_density()

```

This graph is similar but has all the continents filled in.

```{r}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.7)
```


### Exercise 2

Here is a density plot with a lower alpha to better display the data.

```{r plastic-waste-density}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.25)
```

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.

```{r plastic-waste-violin}
# insert code here
```

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-mismanaged}
# insert code here
```

```{r plastic-waste-mismanaged-continent}
# insert code here
```

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r recreate-viz}
# insert code here
```


