---
title: "Lab 02 - Plastic waste"
author: Ben Waggener
date: 01/21/2025
output: github_document
---

## Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv("data/plastic-waste.csv")
```

## Exercises

### Exercise 1

Here is the plastic-waste-continent data plotted as a frequency graph 

```{r plastic-waste-continent}
ggplot(data = plastic_waste, aes(x = plastic_waste_per_cap)) +
  geom_histogram(binwidth = 0.2)
```

Here is the dataset filtered so that it only shows countries with plastic waste per capita that is greater than 3.5kg per person
```{r filtered_plasticwastecontinent}

plastic_waste %>%
  filter(plastic_waste_per_cap > 3.5)
```
This shows that the outlier from the previous graph is Trinidad and Tobago. This means that this country has a much higher rate of plastic waste per capita compared to the other countries in this dataset. Quickly looking online, it seems that this is not a "data error" and that much of this is caused by "inadequate waste management" (according to Forbes)


Below shows each continent as a density plot.

```{r desnsityplotbycontinent}

ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent
  )
) +
  geom_density()

```

This graph is similar but has all the continents filled in.

```{r}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.7)
```


### Exercise 2


"2.1. Recreate the density plots above using a different (lower) alpha level that works better for displaying the density curves for all continents."

```{r plastic-waste-density}
ggplot(
  data = plastic_waste,
  mapping = aes(
    x = plastic_waste_per_cap,
    color = continent,
    fill = continent
  )
) +
  geom_density(alpha = 0.25)
```


"2.2. Explain why we defined the color and fill of the curves by mapping aesthetics of the plot but we defined the alpha level as a characteristic of the plotting geom."

'aes' is used to communicate to R what data to use whereas 'geom' says how to display that data. In the graph above, the x axis displays plastic waste per capita while color and fill display continent. These are under the 'aes' function in R. Alpha refers to the color transparency which is a quality of the data visualization rather than identifying the data itself.


🧶 ✅ ⬆️ Now is a good time to knit your document and commit and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.

And yet another way to visualize this relationship is using side-by-side box plots."


Here is a density plot with a lower alpha to better display the data.



### Exercise 3

Remove this text, and add your answer for Exercise 3 here.

```{r plastic-waste-violin}
# insert code here
```

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-mismanaged}
# insert code here
```

```{r plastic-waste-mismanaged-continent}
# insert code here
```

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r recreate-viz}
# insert code here
```


